---
title: "Visualization for RDD"
output: html_notebook
---

Summary. The difference of y_hats at the cutoff (i.e., Y_hat(treated cases) - Y_hat(counterfactually treated cases)) only agree with estimated effect when the RDD is sharp w/wo covariates. The preceding difference disagrees with the estimate even when there is no covariates. I think this is because in a Fuzzy design: Y_hat ~ Z_hat + (X-c)Tr + (X-c)(Tr-1), Z_hat ≠ Tr and hence the coefficient of Z_hat is not the discontinuity of fitted lines at c. Alterntively, if I predict Y_hat uisng Y_hat ~ Tr + (X-c)Tr + (X-c)(Tr-1), the difference at c will agree with the estimate, but the plot looks ugly.

Fit a RDD, cutpoint == 85 is a sharp design.
```{r model}
library(rddapp)
dat = read.csv('./dat_s.csv')
shp_w_cov = RDestimate(SBIQ48 ~ MOMWAIS0 + DC_TRT|SBIQ24+SBIQ36 , data = dat, 
  cutpoint = 85, model = TRUE, frame = TRUE)
shp_wo_cov = RDestimate(SBIQ48 ~ MOMWAIS0 + DC_TRT , data = dat, 
  cutpoint = 85, model = TRUE, frame = TRUE)
fzy_w_cov = RDestimate(SBIQ48 ~ MOMWAIS0 + DC_TRT|SBIQ24+SBIQ36 , data = dat, 
  cutpoint = 87, model = TRUE, frame = TRUE)
fzy_wo_cov = RDestimate(SBIQ48 ~ MOMWAIS0 + DC_TRT , data = dat, 
  cutpoint = 87, model = TRUE, frame = TRUE)
```

Prepareing model data
```{r model.data}
RDpredict(shp_w_cov, plot = T)
```


Predictions of parametric models: Y_hat = Z_hat + (X-C)(1-Tr) + (X-C)Tr
```{r predict}
## PREDICTION FOR PARAMETRIC MODEL AT UNIQUE X (AND GRAND MEANS OF COVARIATES) ##   
df_p = subset(df, select = -c(Y, Z)) 

# ADD COUNTERFACTUALS AT CUTOFF
df_p$cf = 0

df_p_cut = if(cut %in% df_p$X) {
  within(subset(df_p, X==cut), {
    Tr = 0
    cf = 1
  })
} else {
  data.frame(X = rep(cut,2), 
    rbind(sapply(df_p[covs], mean, na.rm=T), sapply(df_p[covs], mean, na.rm=T)),
    X3 = c(0,0),  X2 = c(0,0),  X1 = c(0,0), Tr = c(0,1),  cf = 1:0)
}

df_p = rbind(df_p, setNames(df_p_cut, names(df_p)))

# STAGE 1. PREDICT Z_HAT (AT ACTUAL COVS)
df_p = within(df_p, {
  Z1 = drop(as.matrix(cbind(1, Tr, X1*(1-Tr), X1*Tr, df_p[covs])) %*%
      coef(rd$model$firststage[[1]]))
  Z2 = drop(as.matrix(cbind(1, Tr, X1*(1-Tr), X2*(1-Tr), X1*Tr, X2*Tr, df_p[covs])) %*%
      coef(rd$model$firststage[[2]]))
  Z3 = drop(as.matrix(cbind(1, Tr, X1*(1-Tr), X2*(1-Tr), X3*(1-Tr), X1*Tr, X2*Tr, X3*Tr,  df_p[covs])) %*% 
      coef(rd$model$firststage[[3]]))
})

# SET COVS AT GRAND MEAN
if(length(covs)>0){
  cov_m = sapply(subset(rd$frame,select=covs), mean, na.rm = T)
  df_p = cbind(subset(df_p, select = -which(names(df_p) %in% covs)), as.list(cov_m))
} 


# STAGE 2. PREDICT Y_HAT (MEANS OF COVS)
df_p = within(df_p, {
  
  Y1 = as.matrix(cbind(1, Z1, X1*(1-Tr), X1*Tr, df_p[covs])) %*% coef(rd$model$iv[[1]])
  Y2 = as.matrix(cbind(1, Z2, X1*(1-Tr), X2*(1-Tr),X1*Tr, X2*Tr,  df_p[covs])) %*% coef(rd$model$iv[[2]])
  Y3 = as.matrix(cbind(1, Z3, X1*(1-Tr), X2*(1-Tr), X3*(1-Tr), X1*Tr, X2*Tr, X3*Tr,  df_p[covs])) %*% coef(rd$model$iv[[3]])
  
  VAR1 = rowSums(as.matrix(cbind(1, Z1, X1*(1-Tr), X1*Tr,  df_p[covs])) %*% vcov(rd$model$iv[[1]]) * as.matrix(cbind(1, Z1, X1*(1-Tr), X1*Tr,df_p[covs])))
  VAR2 = rowSums(as.matrix(cbind(1, Z2, X1*(1-Tr), X2*(1-Tr), X1*Tr, X2*Tr,df_p[covs])) %*% vcov(rd$model$iv[[2]]) * as.matrix(cbind(1, Z2, X1*(1-Tr), X2*(1-Tr),X1*Tr, X2*Tr,  df_p[covs])))
  VAR3 = rowSums(as.matrix(cbind(1, Z3, X1*(1-Tr), X2*(1-Tr), X3*(1-Tr), X1*Tr, X2*Tr, X3*Tr,  df_p[covs])) %*% vcov(rd$model$iv[[3]]) * as.matrix(cbind(1, Z3, X1*(1-Tr), X2*(1-Tr), X3*(1-Tr),X1*Tr, X2*Tr, X3*Tr,   df_p[covs])))
  
  Y1l = Y1 + qnorm((1-plevel)/2) * sqrt(VAR1)
  Y1u = Y1 - qnorm((1-plevel)/2) * sqrt(VAR1)
  Y2l = Y2 + qnorm((1-plevel)/2) * sqrt(VAR2)
  Y2u = Y2 - qnorm((1-plevel)/2) * sqrt(VAR2)
  Y3l = Y3 + qnorm((1-plevel)/2) * sqrt(VAR3)
  Y3u = Y3 - qnorm((1-plevel)/2) * sqrt(VAR3)
})

df_p=df_p[order(df_p$X),]
```


Plot predictions
```{r plot, fig.height=8,fig.width=8}
## TEST PLOT
plot.new()
plot.window(xlim = range(rd$frame$X), ylim = range(rd$frame$Y))
points(Y ~ X, subset(rd$frame, Z==1), pch=19, cex = .5)
points(Y ~ X, subset(rd$frame, Z==0), pch=1, cex = .5)
abline(v = cut, lty = 3);box()

out=by(df_p, factor(df_p$Tr), function (df){
  points(Y1 ~ X ,df, pch=1, col = 'blue', type = 'b')
  points(Y2 ~ X ,df, pch=2, col = 'green', type = 'b')
  points(Y3 ~ X ,df, pch=3, col = 'red', type = 'b')
  # points(Y.o~X, pred, pch=20, col = 'black', type = 'l')
})

```

Compare predictions at cutoff with the estimates
```{r effect}
rbind(
  RDestimate = rd$est[1:3], 
  Pred_diff = subset(df_p, X==cut & Tr == 1, select = Y1:Y3) - subset(df_p, X==cut & Tr == 0,select = Y1:Y3)
) 

```

## Alternative 1 ##

Predictions of parametric models: Y_hat = Z_hat + (X-C)(1-Tr) + (X-C)Tr
```{r predict1}
## PREDICTION FOR PARAMETRIC MODEL AT UNIQUE X (AND GRAND MEANS OF COVARIATES) ##   
df_p = subset(df, select = -c(Y, Z)) 

# ADD COUNTERFACTUALS AT CUTOFF
df_p$cf = 0

df_p1_cut = if(cut %in% df_p$X) {
  within(subset(df_p, X==cut), {
    Tr = 0
    cf = 1
  })
} else {
  data.frame(X = rep(cut,2), 
    rbind(sapply(df_p[covs], mean, na.rm=T), sapply(df_p[covs], mean, na.rm=T)),
    X3 = c(0,0),  X2 = c(0,0),  X1 = c(0,0), Tr = c(0,1),  cf = 1:0)
}

df_p = rbind(df_p, setNames(df_p1_cut, names(df_p)))

# STAGE 1. PREDICT Z_HAT (AT ACTUAL COVS)
df_p = within(df_p, {
  Z1 = drop(as.matrix(cbind(1, Tr, X1*(1-Tr), X1*Tr, df_p[covs])) %*%
      coef(rd$model$firststage[[1]]))
  Z2 = drop(as.matrix(cbind(1, Tr, X1*(1-Tr), X2*(1-Tr), X1*Tr, X2*Tr, df_p[covs])) %*%
      coef(rd$model$firststage[[2]]))
  Z3 = drop(as.matrix(cbind(1, Tr, X1*(1-Tr), X2*(1-Tr), X3*(1-Tr), X1*Tr, X2*Tr, X3*Tr,  df_p[covs])) %*% 
      coef(rd$model$firststage[[3]]))
})

# SET COVS AT GRAND MEAN
if(length(covs)>0){
  cov_m = sapply(subset(rd$frame,select=covs), mean, na.rm = T)
  df_p = cbind(subset(df_p, select = -which(names(df_p) %in% covs)), as.list(cov_m))
} 

# SET Z_hats AT MEANS IN EACH SIDE

Z_m = aggregate(subset(df_p,select=Z1:Z3), by = list(Tr=df_p$Tr), mean, na.rm = T)
df_p = merge(subset(df_p, select = -c(Z1:Z3)), Z_m, by = 'Tr')

# STAGE 2. PREDICT Y_HAT (MEANS OF COVS)
df_p = within(df_p, {
  
  Y1 = as.matrix(cbind(1, Z1, X1*(1-Tr), X1*Tr, df_p[covs])) %*% coef(rd$model$iv[[1]])
  Y2 = as.matrix(cbind(1, Z2, X1*(1-Tr), X2*(1-Tr),X1*Tr, X2*Tr,  df_p[covs])) %*% coef(rd$model$iv[[2]])
  Y3 = as.matrix(cbind(1, Z3, X1*(1-Tr), X2*(1-Tr), X3*(1-Tr), X1*Tr, X2*Tr, X3*Tr,  df_p[covs])) %*% coef(rd$model$iv[[3]])
  
  VAR1 = rowSums(as.matrix(cbind(1, Z1, X1*(1-Tr), X1*Tr,  df_p[covs])) %*% vcov(rd$model$iv[[1]]) * as.matrix(cbind(1, Z1, X1*(1-Tr), X1*Tr,df_p[covs])))
  VAR2 = rowSums(as.matrix(cbind(1, Z2, X1*(1-Tr), X2*(1-Tr), X1*Tr, X2*Tr,df_p[covs])) %*% vcov(rd$model$iv[[2]]) * as.matrix(cbind(1, Z2, X1*(1-Tr), X2*(1-Tr),X1*Tr, X2*Tr,  df_p[covs])))
  VAR3 = rowSums(as.matrix(cbind(1, Z3, X1*(1-Tr), X2*(1-Tr), X3*(1-Tr), X1*Tr, X2*Tr, X3*Tr,  df_p[covs])) %*% vcov(rd$model$iv[[3]]) * as.matrix(cbind(1, Z3, X1*(1-Tr), X2*(1-Tr), X3*(1-Tr),X1*Tr, X2*Tr, X3*Tr,   df_p[covs])))
  
  Y1l = Y1 + qnorm((1-plevel)/2) * sqrt(VAR1)
  Y1u = Y1 - qnorm((1-plevel)/2) * sqrt(VAR1)
  Y2l = Y2 + qnorm((1-plevel)/2) * sqrt(VAR2)
  Y2u = Y2 - qnorm((1-plevel)/2) * sqrt(VAR2)
  Y3l = Y3 + qnorm((1-plevel)/2) * sqrt(VAR3)
  Y3u = Y3 - qnorm((1-plevel)/2) * sqrt(VAR3)
})

df_p=df_p[order(df_p$X),]
```


Plot predictions
```{r plot1, fig.height=8,fig.width=8}
## TEST PLOT
plot.new()
plot.window(xlim = range(rd$frame$X), ylim = range(rd$frame$Y, df_p[c('Y1','Y2','Y3')]))
points(Y ~ X, subset(rd$frame, Z==1), pch=19, cex = .5)
points(Y ~ X, subset(rd$frame, Z==0), pch=1, cex = .5)
abline(v = cut, lty = 3);box()

out=by(df_p, factor(df_p$Tr), function (df){
  points(Y1 ~ X ,df, pch=1, col = 'blue', type = 'b')
  points(Y2 ~ X ,df, pch=2, col = 'green', type = 'b')
  points(Y3 ~ X ,df, pch=3, col = 'red', type = 'b')
  # points(Y.o~X, pred, pch=20, col = 'black', type = 'l')
})

```

Compare predictions at cutoff with the estimates
```{r effect1}
rbind(
  RDestimate = rd$est[1:3], 
  Pred_diff = subset(df_p, X==cut & Tr == 1, select = Y1:Y3) - subset(df_p, X==cut & Tr == 0,select = Y1:Y3)
) 

```
## Alternative 2 ##

Predictions of parametric models: Y_hat = Z_hat + (X-C)(1-Tr) + (X-C)Tr
```{r predict2}
## PREDICTION FOR PARAMETRIC MODEL AT UNIQUE X (AND GRAND MEANS OF COVARIATES) ##   
df_p = subset(df, select = -c(Y, Z)) 

# ADD COUNTERFACTUALS AT CUTOFF
df_p$cf = 0

df_p1_cut = if(cut %in% df_p$X) {
  within(subset(df_p, X==cut), {
    Tr = 0
    cf = 1
  })
} else {
  data.frame(X = rep(cut,2), 
    rbind(sapply(df_p[covs], mean, na.rm=T), sapply(df_p[covs], mean, na.rm=T)),
    X3 = c(0,0),  X2 = c(0,0),  X1 = c(0,0), Tr = c(0,1),  cf = 1:0)
}

df_p = rbind(df_p, setNames(df_p1_cut, names(df_p)))

# STAGE 1. PREDICT Z_HAT (AT ACTUAL COVS)
df_p = within(df_p, {
  Z1 = drop(as.matrix(cbind(1, Tr, X1*(1-Tr), X1*Tr, df_p[covs])) %*%
      coef(rd$model$firststage[[1]]))
  Z2 = drop(as.matrix(cbind(1, Tr, X1*(1-Tr), X2*(1-Tr), X1*Tr, X2*Tr, df_p[covs])) %*%
      coef(rd$model$firststage[[2]]))
  Z3 = drop(as.matrix(cbind(1, Tr, X1*(1-Tr), X2*(1-Tr), X3*(1-Tr), X1*Tr, X2*Tr, X3*Tr,  df_p[covs])) %*% 
      coef(rd$model$firststage[[3]]))
})

# SET COVS AT GRAND MEAN
if(length(covs)>0){
  cov_m = sapply(subset(rd$frame,select=covs), mean, na.rm = T)
  df_p = cbind(subset(df_p, select = -which(names(df_p) %in% covs)), as.list(cov_m))
} 

# SET Z_hats AT 1 or 0 IN EACH SIDE

Z_m = aggregate(subset(df_p,select=Z1:Z3), by = list(Tr=df_p$Tr), mean, na.rm = T)
Z_m = round(Z_m)
df_p = merge(subset(df_p, select = -c(Z1:Z3)), Z_m, by = 'Tr')

# STAGE 2. PREDICT Y_HAT (MEANS OF COVS)
df_p = within(df_p, {
  
  Y1 = as.matrix(cbind(1, Z1, X1*(1-Tr), X1*Tr, df_p[covs])) %*% coef(rd$model$iv[[1]])
  Y2 = as.matrix(cbind(1, Z2, X1*(1-Tr), X2*(1-Tr),X1*Tr, X2*Tr,  df_p[covs])) %*% coef(rd$model$iv[[2]])
  Y3 = as.matrix(cbind(1, Z3, X1*(1-Tr), X2*(1-Tr), X3*(1-Tr), X1*Tr, X2*Tr, X3*Tr,  df_p[covs])) %*% coef(rd$model$iv[[3]])
  
  VAR1 = rowSums(as.matrix(cbind(1, Z1, X1*(1-Tr), X1*Tr,  df_p[covs])) %*% vcov(rd$model$iv[[1]]) * as.matrix(cbind(1, Z1, X1*(1-Tr), X1*Tr,df_p[covs])))
  VAR2 = rowSums(as.matrix(cbind(1, Z2, X1*(1-Tr), X2*(1-Tr), X1*Tr, X2*Tr,df_p[covs])) %*% vcov(rd$model$iv[[2]]) * as.matrix(cbind(1, Z2, X1*(1-Tr), X2*(1-Tr),X1*Tr, X2*Tr,  df_p[covs])))
  VAR3 = rowSums(as.matrix(cbind(1, Z3, X1*(1-Tr), X2*(1-Tr), X3*(1-Tr), X1*Tr, X2*Tr, X3*Tr,  df_p[covs])) %*% vcov(rd$model$iv[[3]]) * as.matrix(cbind(1, Z3, X1*(1-Tr), X2*(1-Tr), X3*(1-Tr),X1*Tr, X2*Tr, X3*Tr,   df_p[covs])))
  
  Y1l = Y1 + qnorm((1-plevel)/2) * sqrt(VAR1)
  Y1u = Y1 - qnorm((1-plevel)/2) * sqrt(VAR1)
  Y2l = Y2 + qnorm((1-plevel)/2) * sqrt(VAR2)
  Y2u = Y2 - qnorm((1-plevel)/2) * sqrt(VAR2)
  Y3l = Y3 + qnorm((1-plevel)/2) * sqrt(VAR3)
  Y3u = Y3 - qnorm((1-plevel)/2) * sqrt(VAR3)
})

df_p=df_p[order(df_p$X),]
```

Plot predictions
```{r plot2, fig.height=8,fig.width=8}
## TEST PLOT
plot.new()
plot.window(xlim = range(rd$frame$X), ylim = range(rd$frame$Y, df_p[c('Y1','Y2','Y3')]))
points(Y ~ X, subset(rd$frame, Z==1), pch=19, cex = .5)
points(Y ~ X, subset(rd$frame, Z==0), pch=1, cex = .5)
abline(v = cut, lty = 3);box()

out=by(df_p, factor(df_p$Tr), function (df){
  points(Y1 ~ X ,df, pch=1, col = 'blue', type = 'b')
  points(Y2 ~ X ,df, pch=2, col = 'green', type = 'b')
  points(Y3 ~ X ,df, pch=3, col = 'red', type = 'b')
  # points(Y.o~X, pred, pch=20, col = 'black', type = 'l')
})

```

Compare predictions at cutoff with the estimates
```{r effect2}
rbind(
  RDestimate = rd$est[1:3], 
  Pred_diff = subset(df_p, X==cut & Tr == 1, select = Y1:Y3) - subset(df_p, X==cut & Tr == 0,select = Y1:Y3)
) 

```
